### 2013-02-08 Chapters/Chapter2/Section_signals.tex
Diagrams for sensor processing

% Section Signal Pre-processing
% !TEX root = ../../main.tex

\section{Signal pro-processing and sensor fusion for timed patterns}
  Describe how data is collected and processed, the form of the data (continuous to discrete, amplitude, energy, frequencies, etc).
  Include feature extraction, like PCA, FFT, energy, mean, but discuss it further in subsection.
  Only the concepts will be explained and the scheme of the computations (or simply formulas), no precise implementations.
  Mention only shortly with meaning, no further explanation.

    \subsubsection{Signal fusion}
    What, why and how.

    \subsubsection{Signal smoothing}
    What, why and how.

    \subsubsection{Feature extraction}
    What, why and how.