### 2013-05-08 Chapters/old_classification/Chapter1/literature_review_nonparametric_clustering.tex
Create outlines for new focus: temporal segmentation

%
% !TEX root = ../../main.tex

\subsection{Non-parametric clustering methods}\label{sec:lit_review_nonparametric}
Many clustering methods rely on either a known number of expected clusters on implement an iterative approach to discover the ideal number of clusters.
In non-parametric approaches the number of clusters is a priori unknown and this discovery is intrinsic in the method, besides the labeling of the data.
This subsection will review the applied non-parametric methods in the field of clustering, where possible of time series.