### 2013-05-14 Chapters/Chapter2/cusum_glr.tex
Add introduction Chapter 2

% !TEX root = ../../main.tex
\section{CUSUM and GLR}\label{cusum-glr}

Detection of variance: \cite{inclan1994use}.