### 2013-05-22 Chapters/Chapter2/cusum_glr.tex
Working on SVM

% !TEX root = ../../main.tex
\section{CUSUM and GLR}\label{cusum-glr}

*** read book Basseville \cite{basseville1993detection} ***

\subsection{CUSUM}
Non-Bayesian change detection algorithm (thus: no prior distribution believe available for the change time).

The CUSUM (cumulative sum) method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).

Primary for detection of mean shift.

The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}.

Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance \cite{inclan1994use}

\subsection{GLR}
Also: maximum-likelihood estimation. ``When applied to a dataset and a given statistical model, it provides estimates for the model's parameters.''