### 2013-07-03 glossary-acronyms.tex

% !TEX root = main.tex

% A C R O N Y M S

\newacronym{cusum}{CUSUM}{Cumulative Sum}

\newacronym{icss}{ICSS}{Iterated Cumulative Sums of Squares}

\newacronym{mosum}{MOSUM}{Moving Sum}

\newacronym{pdf}{PDF}{Probability Density Function}

\newacronym{svdd}{SVDD}{Support Vector Data Description}

\newacronym{svm}{SVM}{Support Vector Machine}