### 2013-07-03 Chapters/Chapter3/outlier_detection.tex
chapter 3: svm outlier

% !TEX root = ../../main.tex
\section{Outlier and change detection}\label{sec:outlier_and_change_detection}