### 2013-08-26 Appendices/AppendixD-planning.tex
WIP

% !TEX root = ../main.tex
% Appendix D

\chapter{Planning} % Main appendix title

\label{AppendixD}

\lhead{Appendix D. \emph{Planning}}

\begin{center}
  \begin{tabulary}{\textwidth}{|c|L|L|c|}
    \hline
    Chapter & Title & Extra & Deadline \\
    \hline \hline
    1 & Introduction, abstract & Write at last & 15/10/2013 \\
    \hline
    2 & Literature review & Partially writting. Maybe focus more on SVM and quick look at other segmentation techniques & 15/9/2013 \\
    \hline
    3 & (Change detection By) One-Class Support Vector Machines & Mix of current chapter and blog post & 1/10/2013 \\
    \hline
    4 & Proposed method & Propose new method, or just setup of testing the OC-SVMs with accelerometer data & Date \\
    \hline
    5 & \textbf{Results} & Test with known and public databases of accelerometer data. For this, two implementations (in Matlab?) are needed & 15/8/2013 \\
    \hline
    6 & Real-world applications & Test with own labeled accelerometer data & 1/9/2013 \\
    \hline
    7 & Conclusion & Write last, together with introduction & 15/10/2013 \\
    \hline
  \end{tabulary}
\end{center}