### 2013-09-23 Chapters/Chapter4/change_indication.tex
Writing on method

% !TEX root = ../../main.tex
\section{Change Indication}\label{sec:method_change_indication}