### 2013-10-08 Chapters/Chapter3.tex
Add note on paper

% !TEX root = ../main.tex
% Chapter 3

\chapter{Change detection by Support Vector Machines}

\label{Chapter3} % For referencing the chapter elsewhere, use \ref{Chapter3}

\lhead{Chapter 3. \emph{Change detection by Support Vector Machines}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Content follows blogpost, but more formal
  \item Explain the connection between change detection and \{outlier detection, density estimation, \etc\}
  \item Explain change detection in relation with ``M-Estimators''.
  \item Explain why and how SVM can be used for outlier detection, and thus for change detection in time series
  \item Explain options when using SVM, such as RBF kernel
  \item Relate characteristics of (accelerometer) data to the used (RBF?) kernel
  \item Define some quality metrics --> this should be in a different chapter (About experiments)
  \item Note: try to avoid the perspective of \emph{density estimation}. Instead: \emph{data description} or \emph{boundary description}.
\end{itemize}

*** TODO: ***
Follow methodology of ``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}.
It creates a two-stage process of first searching for outliers, and then using the ``outlier-score'' to find (sudden) change points, by a weighted average of a moving window.
That eventual score can be thresholded (as in the paper) or processed with something like CUSUM (proposal).
Looks like my proposed methods, in that is combines outliers and gives a score to change points.

\input{./Chapters/Chapter3/outlier_detection}
\input{./Chapters/Chapter3/one_class_svm}
\input{./Chapters/Chapter3/svm_parameters}
\input{./Chapters/Chapter3/data_characteristics}
\input{./Chapters/Chapter3/quality_metrics}