### 2013-10-23 Chapters/Chapter3/one_class_classification.tex
More on OCC

% !TEX root = ../../main.tex
\section{One-Class Classification}\label{sec:one_class_classification}

*** TODO: bridge that explains why \gls{occ} is interesting ***

\subsection{Problem formulation}\label{subsec:occ-problem-formulation}
*** Introduce problem of \gls{occ}. Relate to two-class classification. ***

The problem of \gls{occ} is closely related to the (traditional) two-class classification situation\footnote{Two-class problems are considered as to be the basic problem, since multi-class classification problems can be decomposed into multiple two-class problems \cite{fukunaga1990introduction}.}.
In the case of traditional classification algorithms, the problem is to classify an unknown object to one of the pre-defined categories.
Every object $i$ is represented as a vector $\vectorsym{x}_i = (x_{i,1},\dots,x_{i,d}), x_{i,j} \in \mathbb{R}$ with $d$ real-valued measurements.
Using this notation, an object $\vectorsym{x}_i$ thus represents one point in a feature space: $\vectorsym{x}_i \in \mathcal{X} = \mathbb{R}^d$.
The two classes of objects, $\omega_1$ and $\omega_2$, are labeled $-1$ and $+1$ respectively.
The objects with a label $y_i \in \{-1, +1\}$ attached are in the training set $\mathcal{X}^{tr}$ (note that it can be both positive and negative example objects).
This problem is solved by determining a decision boundary in the feature space of the data objects and label the new data object based on the location relative to this boundary.
This is expressed by the decision function $y = f(\vectorsym{x})$:
\begin{equation}\label{eq:decision_function_classification}
  f: \mathbb{R}^d \rightarrow \{-1, +1\}
\end{equation}
In case of the \gls{occ} problem, only one class (often referred as the target class, or positive examples) of training data is used to create a decision boundary.
The goal is to determine whether a new data object belongs to the target class or to the unknown class and thus is an outlier.
One could argue that this problem is equal to the traditional two-class problem by considering all other classes as negative examples, although there are important differences.
In pure \gls{occ} problems there are no negative example objects available.
This could be because the acquisition of these examples is very expensive, or because there are only examples of the `normal' state of the system and the goal is to detect `abnormal' states.
Since the algorithm's goal is to differentiate between normal and abnormal objects (relative to the training objects), \gls{occ} is often called outlier, novelty or anomaly detection, depending on the origin of the problem to which the algorithm is applied\footnote{The term \acrlong{occ} originates from Moya \etal \cite{moya1993one}.}.
The difference between two and one-class classification and the consequence for outlier objects is illustrated in Figure \ref{fig:two-vs-one-classification}.
In the two-class classification problem the object \textbf{o} will be member of the \textbf{-1} class whilst the \gls{occ} problem will label it as an outlier.
In \cite{tax2001one} a more detailed analysis of the \gls{occ} is given.

\begin{figure}
  \centering
    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/two-vs-one-classification.pdf}
  \caption[Difference between two and one-class classification]{This plot shows the difference between two and one-class classification. The solid line indicates a possible decision boundary between the \textbf{+1} and \textbf{-1} example objects. The dashed circle indicates the closed boundary around all the data objects. In the first type the object \textbf{o} is considered to be member of the \textbf{-1}-class, whilst in the latter (\gls{occ}) formulation it is an outlier.}
  \label{fig:two-vs-one-classification}
\end{figure}

The \gls{occ} algorithms have been applied to a wide range of applications.
The first is, obviously, outlier detection of objects which do not resemble the bulk of the training data.
It can be a goal by itself and can be used as a filtering mechanism for other data processing methods.
Often methods that rely on data-characteristics are sensitive for remote regions in the data set.
Using \gls{occ} these remote regions can be removed from the data set.
A second application is for the problem as described above, in which only data from a single target class is available.
When this problems originates from \eg a monitoring process, the \gls{occ} is able to recognize abnormal system states, without the need to create (or simulate) faulty states beforehand.
The final possible application given by Tax \cite{tax2001one} is the comparison of two data sets.
By constructing a \gls{occ}-classifier using a training data set, one can compare that set to a new data set.
This will result in a similarity-measure.
It is easy to see here a relation with similiary measures over \glspl{pdf} and the Kullback-Leibler divergence.
This application will be further discussed in Section \ref{subsec:change-detection-with-occ}.

\subsection{One-Class Classification methods}\label{subsec:occ-methods}
In his thesis \cite{tax2001one}, Tax orders a collection of \gls{occ}-methods into three categories, visually represented in Figure \ref{fig:occ-methods}.
The first category consists of methods that estimate the density of the training data and set a threshold on this density.
Among those are Gaussian models, \glspl{gmm} and Parzen density estimators.
In order to get good generalization results with these methods, the dimensionality of the data and the complexity of the density need to be restricted.
This can cause a large bias on the data.
When a good probability model is assumed, these methods work very well, since when one threshold is optimized, a minimum volume is automatically found for the given probability density model \cite{tax2001one}.

Boundary methods are based on Vapnik's principle\footnote{With a limited amount of data available, on should avoid solving a more general problem as an intermediate step to solve the original problem \cite{vapnik1998statistical}.} which imply in this case that estimating the complete data density for a \gls{occ} may be too complex, in case one is only interested in the closed boundary.
Examples of methods that focus on the boundary (a direct threshold) of the training data distribution are K-centers, Nearest-neighborhood and \gls{svdd}.
Especially the latter has a strong bias towards minimal volume solutions.
These type of methods are sensitive to scaling of features, since they rely on a well-defined distance measure.
The number of objects that is required is smaller than in the case of density methods.
The boundary method \gls{svdd}, constructed by Tax and which has shown good performance \cite{khan2010survey}, will be further discussed in Section \ref{subsec:oc-svm-svdd}.

Reconstruction methods take a different approach.
Instead of focusing on classification of the data and thus on the discriminating features of data objects, they model the data.
This results in a compact representation of the target data and for any object a reconstruction error can be defined.
Outliers are data objects with a high reconstruction error, since they are worse represented by the constructed model.
Examples of reconstruction methods are K-means, \gls{pca} and some neural network implementations.

\begin{figure}
  \centering
    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/occ_methods.pdf}
  \caption[\gls{occ} methods]{Overview of \gls{occ} methods categorized in Density Estimators, Reconstruction methods and Boundary methods. This categorization follows the definition of Tax in \cite{tax2001one}.}
  \label{fig:occ-methods}
\end{figure}

In \cite{khan2010survey,noumir2012simple} an overview of applications for \gls{occ}-algorithms, and explictly for \gls{svm}-based methods (such as \gls{svdd} and \gls{nu-svm}), is given.
It shows succesful applications for, amonst others, problems in the field of Handwritten Digit Recognition, Face Recognition Applications, Spam Detection and Anomaly Detection \cite{li2003improving,perdisci2006using}.
As discussed in Section \ref{sec:outlier_detection}, this can be used for change detection.

\subsection{Change detection with \gls{occ}-methods}\label{subsec:change-detection-with-occ}
*** Write -short- how \gls{occ}-methods can be used for change detection. Relate to previous section on change detection, \ref{sec:outlier_detection} ***

*** TODO: end with bridge to next chapter, \gls{oc-svm} ***

-- Literature --

% ``A survey of recent trends in one class classification'' \cite{khan2010survey}. 25, 2010 \\

``One-class classification by combining density and class probability estimation'' \cite{hempstalk2008one}. 51, 2008 \\

% ``One-class classification'' \cite{tax2001one}. 693, 2001 \\

``The One-Class Classification Approach to Data Description and to Models Applicability Domain'' \cite{baskin2010one}. 19, 2009 \\

``One-class classifier networks for target recognition applications'' \cite{moya1993one}. 70, 1993. Origin of term ``One-Class Classification'' \\