### 2013-10-25 Chapters/Chapter5/artificial_data.tex
Draft, sent to Anne

% !TEX root = ../../main.tex
\section{Artificial Data}\label{sec:artificial_data}
\TODO{Use data from \cite{camci2010change,takeuchi2006unifying}}
In this section we present three data sets as used by \cite{camci2010change,takeuchi2006unifying} to provide for a objetive performance comparison.