### 2013-10-28 Chapters/Chapter4/change_detection.tex

% !TEX root = ../../main.tex
\section{Change Detection}\label{sec:method_change_detection}