### 2013-10-28 Chapters/Chapter4/model_properties.tex

% !TEX root = ../../main.tex
\section{Model Properties}\label{sec:method_model_properties}