### 2013-10-29 glossary-acronyms.tex

% !TEX root = main.tex

% A C R O N Y M S

\newacronym{ar}{AR}{Autoregressive}

\newacronym{aic}{AIC}{Akaike Information Criterion}

\newacronym{bic}{BIC}{Bayesian Information Criterion}

\newacronym{cusum}{CUSUM}{Cumulative Sum}

\newacronym{har}{HAR}{Human Activity Recognition}

\newacronym{gmm}{GMM}{Gaussian Mixture Model}

\newacronym{hmm}{HMM}{Hidden Markov Model}

\newacronym{icss}{ICSS}{Iterated Cumulative Sums of Squares}

\newacronym{iid}{i.i.d.}{independently and identically distributed}

\newacronym{isvdd}{I-SVDD}{Incremental Support Vector Data Description}

\newacronym{kl}{KL}{Kullback-Leibler}

\newacronym{kliep}{KLIEP}{Kullback-Leibler Importance Estimation Procedure}

\newacronym{kkt}{KKT}{Karush–Kuhn–Tucker}

\newacronym{mle}{MLE}{Maximum Likelihood Estimates}

\newacronym{mosum}{MOSUM}{Moving Sum}

\newacronym{nu-svm}{$\nu$-SVM}{$\nu$-Support Vector Machine}

\newacronym{occ}{OCC}{One-Class Classification}

\newacronym{oc-svm}{OC-SVM}{One-Class Support Vector Machine}

\newacronym{pca}{PCA}{Principal Component Analysis}

\newacronym{pdf}{PDF}{Probability Density Function}

\newacronym{qp}{QP}{Quadratic Programming}

\newacronym{rbf}{RBF}{Radial Base Function}

\newacronym{roc}{ROC}{Receiver-Operating Characteristic}

\newacronym{rt}{RT}{Ratio-Thresholding}

\newacronym{sv}{SV}{Support Vector}

\newacronym{svc}{SVC}{Support Vector Classifier}

\newacronym{svcpd}{SVCPD}{Support Vector based Change Detection}

\newacronym{svdd}{SVDD}{Support Vector Data Description}

\newacronym{svm}{SVM}{Support Vector Machine}

\newacronym{swab}{SWAB}{Sliding Window And Bottom-up}