### 2013-11-01 Chapters/old_classification/Chapter1.tex

% !TEX root = ../main.tex
% Chapter 1
%----------------------------------------------------------------------------------------

%perhaps a shortened title
*** [GRAPHICS diagram: sensors --> signal pre-processing --> feature extraction --> analysis [* segmentation, * labeling] --> conclusions]
*** Give context problems, means to reach (e.g. monitoring of patients,  getting conclusions over health and activity).
\chapter{Introduction} % Main chapter title
\label{Chapter1} % For referencing the chapter elsewhere, use~\ref{Chapter1}
\lhead{Chapter 1. \emph{Introduction}} % This is for the header on each page -

\section{Literature review}
This section will provide a literature review per coherent subject of this thesis.

% \input{./Chapters/Chapter1/literature_review_nonparametric_clustering}
\input{./Chapters/Chapter1/lit_rev_segm_and_class}
\input{./Chapters/Chapter1/lit_rev_temp_segm}
\input{./Chapters/Chapter1/paper_list}

\section{Problem statement}

% \section{System architecture}
% In subSection~\ref{sec:lit_review_segmentation} the research in segmentation and classification of human activities will be discussed.
% SubSection~\ref{sec:lit_review_nonparametric} will take a look on applied non-parametric methods in clustering (time series) of data.