### 2013-11-04 Chapters/Chapter4.tex
WIP: Chapter 4

% !TEX root = ../main.tex
% Chapter 4

\chapter{Proposed method}

\label{Chapter4} % For referencing the chapter elsewhere, use~\ref{Chapter4}

\lhead{Chapter 4. \emph{Proposed method}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Based on the problem statement with current research as stated in Chapter~\ref{Chapter3}
  \item Adjust method to needs
  \item Explain using graphs, pseudo-algorithms. Make clear distinction in origin of ideas and why to apply
\end{itemize}

\subsection{Experiment notes}
\begin{itemize}
  \item ICSS/CUSUM is goed in het vinden van variance changes. Niet goed in mean veranderingen.
  \item Problemen met SVM methode: ook na de verandering blijft de threshold nog stijgen;
  \item geprobeerd direct model-reset te doen na change point, maar dan kan je achteraf niet meer verder analyseren.
  \item Veranderingen in mean geven ``extra'' change point wanneer alle data weer in het window zit. Zie~\ref{fig:jumping_mean_ratios} op de 50-tallen.
\end{itemize}

\begin{description}
  \item[Data gathering] Explain data gathering methods. Refer to chapter 5 and 6 for articifial and real-world details.
  \item[Model construction] Explain SVDD model construction and updating.
  \item[Model properties] Explain the SVDD properties that are used to calculate the change indication.
  \item[Change detection] Explain the interpretation of the properties, the possible (modulair) methods to give change indication
\end{description}

\TODO{Write chapter introduction/outline. Give overview of method approach and refer to each section for details of that stage.}

\TODO{Puts emphasis on two-stage framework by Takeuchi and Yamanishi \cite{takeuchi2006unifying}.}.

\begin{figure}
  \centering
    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/method_setup_short.pdf}
  \caption[Method setup]{Schematic overview of the change detection method. The first step is the data gathering, described in Section \ref{sec:method_data_gathering}. Ater the pre-processing, the data is used to construct a \gls{svdd} model, as described in Section \ref{sec:method_model_construction}. Section \ref{sec:method_model_features} describes which features of the model are used for the final change indication algorithm, discussed in Section \ref{sec:method_change_detection}.}
  \label{fig:method_overview_short}
\end{figure}

\input{./Chapters/Chapter4/data_gathering}
\input{./Chapters/Chapter4/model_construction}
\input{./Chapters/Chapter4/model_features}
\input{./Chapters/Chapter4/change_detection}

 --- OLD ---
\input{./Chapters/Chapter4/change_indication}
\input{./Chapters/Chapter4/overview}
\input{./Chapters/Chapter4/algorithms}