### 2013-11-07 Chapters/Chapter2.tex
WIP: Working on organization of Chapter 2

% !TEX root = ../main.tex
% Chapter 2

\chapter{Literature review}

\label{Chapter2} % For referencing the chapter elsewhere, use~\ref{Chapter2}

\lhead{Chapter 2. \emph{Literature review}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Literature review about Temporal Segmentation (previous draft was more about classification)
  \item Consider methods for the context of filter-methods for classification
  \item Take a loot at 3-4 different kind of methods for change detection:
    \begin{itemize}
      \item Introduction with a lot of techniques
      \item Explain why look at a few
      \item CUSUM - or other more traditional methods
      \item Density-ratio estimation
      \item Support Vector Machines (?) - if there are more sources about this
      \item (Dimensionality reduction) --> probably not
      \item \emph{Not to much attention to all techniques, focus is on SVM}
    \end{itemize}
  \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
\end{itemize}

\section{Statistical framework}\label{statistical-framework}
Many applications require the detection of time points at which the underlying properties of a system change.
This problem has received a lot of attention in the fields of data mining, etc... \TODO{list and refs}.
Often this problem is formulated in a statistical framework, by inspecting the underlying data generating \gls{pdf} of the time series data.
A change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean and variance.

The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} follows this approach.
It originates from control methods for detection from benchmarks.
This method and some derivatiges are discussed and analyzed in Section~\ref{cusum}.

Many methods rely on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \glspl{pdf}, instead of estimating each \gls{pdf}.
This approach is discussed and analyzed in Section~\ref{density-ratio}.

The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
The method of Camci \cite{camci2010change} follows an other approach within the statistical framework, by using a \gls{svm}.
One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
The method represents the distribution over the data points as a hyper-sphere in a higher dimension using the kernel trick.
A change in the \gls{pdf} is represented by a change in the radius of this sphere.
Section~\ref{sec:svm_in_change_detection} discusses the \gls{svm}-methods.
\TODO{Put emphasis that this is the source of inspiration for the chosen method?}

% The final method under consideration, change detection via (intrinsic) dimensionality reduction, follows a different point of view.
% Opposed to the other discussed methods, dimensionality reduction is framed in the MDL (Minimum Description Length) framework.
% It uses the estimated underlying number of parameters of the time series as a model for change detection.
% Section~\ref{dim-reduction} discuses this method.

-- Nieuwe indeling Chapter 2 --

\begin{description}
  \item[Intro/outline] Introduction and overview of this chapter. Put in the context of temporal classification.
  \item[Temporal segmentation] Give overview of used temporal segmentation methods, move towards a split in model building/dimensionality reduction and change detection methods
  \item[Dimensionality reduction] Explain that \gls{dr} can be useful in data processing. Name a few techniques, make bridge to \gls{oc-svm}.
  \item[Change detection] Describe methods for \emph{direct} change detection. End with notion
    \begin{description}
      \item[Model fitting]
      \item[Cusum]
      \item[Density ratio]
    \end{description}
  \item[Support Vector Machine based methods]

\end{description}

\section{Overview}\label{sec:literature_review_overview}
As described in Chapter~\ref{Chapter1}, the focus of this research is the application of temporal segmentation to accelerometer data.
This is with the assumption that explicit temporal segmentation can aid the process of temporal classification.

\TODO{Give examples of temporal classification techniques}

In recent years a lot is research is performed in the classification of human activities, recorded by on-body sensors.
Especially since smartphones with accelerometer sensors became more available, research has focussed on recordings from these devices.

\TODO{Include highlights from old-classification text}

In the following section of this is chapter we begin with a review of current literature on temporal segmentation.
Many techniques of temporal segmentation rely in change detection algorithms.
In Section~\ref{sec:change_detection_methods} we will give an overview of those kind of algorithms.
An often encountered problem is the \emph{curse of dimensionality}.
To overcome this problem, we look at \gls{dr} methods in Section~\ref{sec:dimensionality_reduction}.
We end this chapter with a review on \gls{svm} based methods and algorithms.
This latter section relates the \gls{oc-svm} methods via \gls{dr} to temporal segmentation.

\input{./Chapters/Chapter2/temporal_segmentation.tex}
\input{./Chapters/Chapter2/change_detection_methods.tex}
\input{./Chapters/Chapter2/dimensionality_reduction.tex}
\input{./Chapters/Chapter2/svm_based.tex}

-- OLD --

%----------------------------------------------------------------------------------------
% CUSUM and GLR
\input{./Chapters/Chapter2/cusum}

%----------------------------------------------------------------------------------------
% Density Ratio Estimation
\input{./Chapters/Chapter2/density_ratio}

%----------------------------------------------------------------------------------------
% Temporal Segmentation
\input{./Chapters/Chapter2/temporal_segmentation_old.tex}

%----------------------------------------------------------------------------------------
% Support Vector Machines
\input{./Chapters/Chapter2/svm}