### 2013-11-07 Chapters/Chapter2/temporal_segmentation.tex

% !TEX root = ../../main.tex
\section{Temporal Segmentation}\label{sec:temporal_segmentation}

Give overview of used temporal segmentation methods. Problem is that it is not really well-defined.