### 2013-11-07 Chapters/Chapter4/data_gathering.tex
WIP

% !TEX root = ../../main.tex
\section{Data Gathering}\label{sec:method_data_gathering}
In this section we briefly discuss the different data gathering methods used for the change detection algorithms and experiments.
Section~\ref{subsec:data_gathering_artificial} reviews the artificial data sets we will use.
In Section~\ref{subsec:data_gathering_real_world} an overview of the real-world data sets used is provided.
Both sections refer to Chapters~\ref{Chapter5} and~\ref{Chapter6} for more details, respectively.

% ------
\subsection{Artificial data}\label{subsec:data_gathering_artificial}
In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the researches on which our method is based.
These are the data sets used by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
Both construct a collection of one-dimensional time series data according to a second order \gls{ar} model:
\begin{equation}
  x_t = a_1 x_{t-1} + a_2 x_{t-2} + \epsilon_t.
\end{equation}
In the different data series the mean and variance of the Gaussian random variable $\epsilon_t$ differs and changes at pre-determined change points.
Using this data set an objective quality measure over the change detection methods can be obtained and compared.
All the used data sets are listed and analyzed in Chapter~\ref{Chapter5}.

% ------
\subsection{Real-world data}\label{subsec:data_gathering_real_world}
In the second type of data sets we apply our method of change detection and temporal segmentation to real-world data sets.
For our setup we record the activities of humans performed both in- and outdoor in an uncontrolled environment.
Activities performed include sitting, standing, walking, running in a straight and curved line, and walking up- and downstairs.
Our method uses the signals from the accelerometer, gravity, and rotation sensors.
These time series data are used to detect change points.
A video recording from the performed activity is used to annotate the time series with real change points.
The discovered change point are compared with these annotated change points to give a subjective quality measure.
In Chapter~\ref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.

For the experiments we used a HTC Sensation XE smartphone as recording device.
The activities were recorded using a free Android application \cite{sensorlogger}.
This application was chosen for its convenient data format of the sensor recording and its regularity of the sampling interval.
Table~\ref{tab:recorded_metrics} lists the recorded metrics.
For our expiments we used the data for the accelerometer, magnatief field and rotation.

\begin{center}\begin{table}
  \begin{tabulary}{\textwidth}{|l|L|c|c|}
    \hline
    Metric & Description & Units of measure & Typical range \\
    \hline \hline
    Accelerometer & Acceleration force along each axis. & $m/s^2$ & $-20$ -- $20$ \\
    \hline
    Gravity & Force of gravity along each axi.s & $m/s^2$ & $-10$ -- $10$\\
    \hline
    Gyroscope & & $rad/s$ & $-15$ -- $15$\\
    \hline
    Light & Light sensitive sensor at the front of the phone. & & $0$ -- $10000$ \\
    \hline
    Linear acceleration & & $m/s^2$ & $-20$ -- $20$ \\
    \hline
    Magnetic field & Geomagnetic field strength along each axis. & $\mu T$ & $-60$ -- $60$ \\
    \hline
    Orientation & & & $-100$ -- $360$ \\
    \hline
    Rotation & & & $-1$ -- $1$\\
    \hline
  \end{tabulary}
  \caption{Measured metrics. The set of axis is always the triple (x, y, z) direction.}
  \label{tab:recorded_metrics}
\end{table}\end{center}