### 2013-11-14 Chapters/Chapter5/results.tex
WIP: artificial data

% !TEX root = ../../main.tex
\section{Results}\label{sec:artificial_data_results}