### 2013-11-18 Chapters/Chapter3.tex
Restructure Chapter 5 graphs of data sets

% !TEX root = ../main.tex
% Chapter 3

\chapter{Change detection by Support Vector Machines}

\label{Chapter3} % For referencing the chapter elsewhere, use~\ref{Chapter3}

\lhead{Chapter 3. \emph{Change detection by Support Vector Machines}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Content follows blogpost, but more formal
  \item Explain the connection between change detection and \{outlier detection, density estimation, \etc\}
  \item Explain change detection in relation with ``M-Estimators''.
  \item Explain why and how SVM can be used for outlier detection, and thus for change detection in time series
  \item Explain options when using SVM, such as RBF kernel
  \item Relate characteristics of (accelerometer) data to the used (RBF?) kernel
  \item Define some quality metrics --> this should be in a different chapter (About experiments)
  \item Note: try to avoid the perspective of \emph{density estimation}. Instead: \emph{data description} or \emph{boundary description}.
\end{itemize}

This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as discussed in \Cref{Chapter4}.
The first sections formulates the problem of change detection and relates it to outlier and novelty detection.
It transforms the traditional problem to change detection for times series data.
In \Cref{sec:one_class_classification} the problem of \acrlong{occ} is discussed and shows a number of implementations.
Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in \Cref{sec:one_class_svm}.
For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in~\ref{sec:sensor_data_characteristics}.
Finally a theoretic set of quality metrics is discussed in \Cref{sec:quality_metrics}.

\input{./Chapters/Chapter3/change_detection_time_series}
\input{./Chapters/Chapter3/one_class_classification}
\input{./Chapters/Chapter3/one_class_svm}
\input{./Chapters/Chapter3/sensor_data_characteristics}
\input{./Chapters/Chapter3/quality_metrics}