### 2013-11-18 Chapters/Chapter5.tex

% !TEX root = ../main.tex
% Chapter 5

\chapter{Artificial data results}

\label{Chapter5} % For referencing the chapter elsewhere, use~\ref{Chapter5}

\lhead{Chapter 5. \emph{Artificial data results}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Compare proposed method with methods of \Cref{Chapter2}
  \item Provide KL-analysis (as Takeuchi does)
  \item Provide plots, tables, graphs, error rates, precision, etc.
  \item Apply to a multiple of data, to compare to previous research - use that data
  \item Give theoretical analysis about performance. Big-O, memory, run-time, precision.
  \item This sections needs programmed implementations of own method and the ones compared
\end{itemize}

\input{./Chapters/Chapter5/artificial_data_generation}

\input{./Chapters/Chapter5/results}