### 2013-11-22 Chapters/Chapter2.tex
'Refactoring'

% !TEX root = ../main.tex
% Chapter 2

\chapter{Literature review}

\label{Chapter2} % For referencing the chapter elsewhere, use~\ref{Chapter2}

\lhead{Chapter 2. \emph{Literature review}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Literature review about Temporal Segmentation (previous draft was more about classification)
  \item Consider methods for the context of filter-methods for classification
  \item Take a loot at 3-4 different kind of methods for change detection:
    \begin{itemize}
      \item Introduction with a lot of techniques
      \item Explain why look at a few
      \item CUSUM - or other more traditional methods
      \item Density-ratio estimation
      \item Support Vector Machines (?) - if there are more sources about this
      \item (Dimensionality reduction) --> probably not
      \item \emph{Not to much attention to all techniques, focus is on SVM}
    \end{itemize}
  \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
\end{itemize}

\section{Statistical framework}\label{statistical-framework}
Many applications require the detection of time points at which the underlying properties of a system change.
This problem has received a lot of attention in the fields of data mining, etc... \TODO{list and refs}.
Often this problem is formulated in a statistical framework, by inspecting the underlying data generating \gls{pdf} of the time series data.
A change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean and variance.

The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} follows this approach.
It originates from control methods for detection from benchmarks.
This method and some derivatiges are discussed and analyzed in \Cref{sec:cusum}.

Many methods rely on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \glspl{pdf}, instead of estimating each \gls{pdf}.
This approach is discussed and analyzed in \Cref{sec:density-ratio}.

The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
The method of Camci \cite{camci2010change} follows an other approach within the statistical framework, by using a \gls{svm}.
One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
The method represents the distribution over the data points as a hyper-sphere in a higher dimension using the kernel trick.
A change in the \gls{pdf} is represented by a change in the radius of this sphere.
\Cref{sec:svm_in_change_detection} discusses the \gls{svm}-methods.
\TODO{Put emphasis that this is the source of inspiration for the chosen method?}

% The final method under consideration, change detection via (intrinsic) dimensionality reduction, follows a different point of view.
% Opposed to the other discussed methods, dimensionality reduction is framed in the MDL (Minimum Description Length) framework.
% It uses the estimated underlying number of parameters of the time series as a model for change detection.
% \Cref{dim-reduction} discuses this method.

-- Nieuwe indeling Chapter 2 --

\begin{description}
  \item[Intro/outline] Introduction and overview of this chapter. Put in the context of temporal classification.
  \item[Temporal segmentation] Give overview of used temporal segmentation methods, move towards a split in model building/dimensionality reduction and change detection methods
  \item[Dimensionality reduction] Explain that \gls{dr} can be useful in data processing. Name a few techniques, make bridge to \gls{oc-svm}.
  \item[Change detection] Describe methods for \emph{direct} change detection. End with notion
    \begin{description}
      \item[Model fitting]
      \item[Cusum]
      \item[Density ratio]
    \end{description}
  \item[Support Vector Machine based methods]

\end{description}

\input{./Chapters/Chapter2/overview.tex}
\input{./Chapters/Chapter2/temporal_segmentation.tex}
\input{./Chapters/Chapter2/change_detection_methods.tex}
% \input{./Chapters/Chapter2/dimensionality_reduction.tex}
\input{./Chapters/Chapter2/svm_based.tex}

-- OLD --

%----------------------------------------------------------------------------------------
% CUSUM and GLR
\input{./Chapters/Chapter2/cusum}

%----------------------------------------------------------------------------------------
% Density Ratio Estimation
\input{./Chapters/Chapter2/density_ratio}

%----------------------------------------------------------------------------------------
% Temporal Segmentation
\input{./Chapters/Chapter2/temporal_segmentation_old.tex}

%----------------------------------------------------------------------------------------
% Support Vector Machines
\input{./Chapters/Chapter2/svm}