### 2013-11-22 Chapters/Chapter3.tex

% !TEX root = ../main.tex
% Chapter 3

\chapter{Change detection by Support Vector Machines}

\label{Chapter3} % For referencing the chapter elsewhere, use~\ref{Chapter3}

\lhead{Chapter 3. \emph{Change detection by Support Vector Machines}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Content follows blogpost, but more formal
  \item Explain the connection between change detection and \{outlier detection, density estimation, \etc\}
  \item Explain change detection in relation with ``M-Estimators''.
  \item Explain why and how SVM can be used for outlier detection, and thus for change detection in time series
  \item Explain options when using SVM, such as RBF kernel
  \item Relate characteristics of (accelerometer) data to the used (RBF?) kernel
  \item Define some quality metrics --> this should be in a different chapter (About experiments)
  \item Note: try to avoid the perspective of \emph{density estimation}. Instead: \emph{data description} or \emph{boundary description}.
\end{itemize}

This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as discussed in \Cref{Chapter4}.
The first sections formulates the problem of change detection and relates it to outlier and novelty detection.
It transforms the traditional problem to change detection for times series data.
In \Cref{sec:one_class_classification} the problem of \acrlong{occ} is discussed and shows a number of implementations.
Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in \Cref{sec:one_class_svm}.
For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in~\ref{sec:sensor_data_characteristics}.
Finally a theoretic set of quality metrics is discussed in \Cref{sec:quality_metrics}.

\TODO{Give this a nice place in this chapter somewhere, on model building}
Segmentation methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci \etal \cite{avci2010activity}:
\begin{itemize}
  \item \textbf{Top-down} methods iteratively divide the signal in segments by splitting at the best location.
  The algorithm starts with two segments and completes when a certain condition is met, such as when an error value or number of segments $k$ is reached.
  These methods process the data points recursively, which results in a complexity of $O(n^2k)$.
  \item \textbf{Bottom-up} methods are the natural complement to top-down methods.
  They start with creating $n/2$ segments and iteratively join adjacent segments while the value of a cost function for that operation is below a certain value.
  Given the average segment length $L$, the complexity of this method is $O(nL)$.
  \item \textbf{Sliding-window} methods are simple and intuitive for online segmenting purposes.
  It starts with a small initial subsequence of the time series.
  New data points are joined in the segment until the fit-error is above a threshold.
  Since the data is only processed very locally, these methods can yield in poor results \cite{keogh2001online}.
  The complexity is equal to the bottom-up approach, $O(nL)$, where $L$ is the average segment length.
  \item \textbf{\acrlong{swab}}, as introduced by Keogh \etal \cite{keogh2001online}, joins the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
  The algorithm processes the data in two stages.
  The first stage is to join new data points in the current segment created by a sliding window, and pass this to a buffer with space for a few segments.
  The buffer then processes the data Bottom-up and returns the first (left-most) segment as final segment.
  Because this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-up.
  It is stated by Keogh \etal that the complexity of \gls{swab} is a small constant factor worse than that of regular Bottom-up.
\end{itemize}
It is clear that for the application of this research sliding-window and preferably \gls{swab}-based algorithms should be considered. \\
--- end ---

\TODO{Put somewhere difference and emphasis of online vs. offline/batch/post-processing}

\input{./Chapters/Chapter3/change_detection_time_series}
\input{./Chapters/Chapter3/one_class_classification}
\input{./Chapters/Chapter3/one_class_svm}
\input{./Chapters/Chapter3/sensor_data_characteristics}
\input{./Chapters/Chapter3/quality_metrics}