### 2013-12-03 Chapters/Chapter2.tex
Cleanup of sections

% !TEX root = ../main.tex
% Chapter 2

\chapter{Literature review}

\label{Chapter2} % For referencing the chapter elsewhere, use~\ref{Chapter2}

\lhead{Chapter 2. \emph{Literature review}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Literature review about Temporal Segmentation (previous draft was more about classification)
  \item Consider methods for the context of filter-methods for classification
  \item Take a loot at 3-4 different kind of methods for change detection:
    \begin{itemize}
      \item Introduction with a lot of techniques
      \item Explain why look at a few
      \item CUSUM - or other more traditional methods
      \item Density-ratio estimation
      \item Support Vector Machines (?) - if there are more sources about this
      \item (Dimensionality reduction) --> probably not
      \item \emph{Not to much attention to all techniques, focus is on SVM}
    \end{itemize}
  \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
\end{itemize}

\section{Statistical framework}\label{statistical-framework}

-- Nieuwe indeling Chapter 2 --

\begin{description}
  \item[Intro/outline] Introduction and overview of this chapter. Put in the context of temporal classification.
  \item[Temporal segmentation] Give overview of used temporal segmentation methods, move towards a split in model building/dimensionality reduction and change detection methods
  \item[Dimensionality reduction] Explain that \gls{dr} can be useful in data processing. Name a few techniques, make bridge to \gls{oc-svm}.
  \item[Change detection] Describe methods for \emph{direct} change detection. End with notion
    \begin{description}
      \item[Model fitting]
      \item[Cusum]
      \item[Density ratio]
    \end{description}
  \item[Support Vector Machine based methods]

\end{description}
\input{./Chapters/Chapter2/overview.tex}
\input{./Chapters/Chapter2/statistical_models.tex}
\input{./Chapters/Chapter2/temporal_segmentation.tex}
\input{./Chapters/Chapter2/change_detection_methods.tex}

-- OLD --

%----------------------------------------------------------------------------------------
% CUSUM and GLR
\input{./Chapters/Chapter2/cusum}

%----------------------------------------------------------------------------------------
% Density Ratio Estimation
\input{./Chapters/Chapter2/density_ratio}

%----------------------------------------------------------------------------------------
% Support Vector Machines
\input{./Chapters/Chapter2/svm}