### 2013-12-12 Chapters/Chapter5.tex
WIP: artificial data set results

% !TEX root = ../main.tex
% Chapter 5

\chapter{Artificial data results}

\label{Chapter5} % For referencing the chapter elsewhere, use~\ref{Chapter5}

\lhead{Chapter 5. \emph{Artificial data results}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Compare proposed method with methods of \Cref{Chapter2}
  \item Provide KL-analysis (as Takeuchi does)
  \item Provide plots, tables, graphs, error rates, precision, etc.
  \item Apply to a multiple of data, to compare to previous research - use that data
  \item Give theoretical analysis about performance. Big-O, memory, run-time, precision.
  \item This sections needs programmed implementations of own method and the ones compared
\end{itemize}

In this chapter we will apply the proposed algorithm to artificial data sets.
For comparison, we use some of the data sets that are used by Camci~\cite{camci2010change} and Takeuchi \etal \cite{takeuchi2006unifying}.
These data sets are generated by Gaussian noise distributions.
In \Cref{sec:artificial_data_generation} the precise generation method and characteristics of the data sets are discussed.
For our final change detection mechanism we use the \gls{rt}-method, also applied by Camci \cite{camci2010change}.
This is a simple and fast method and allows us to mainly focus on the \gls{oc-svm} model properties that should reflect changes in the data.
The results of these two methods are discussed in \Cref{sec:artificial_data_results}.

\input{./Chapters/Chapter5/artificial_data_generation}

\input{./Chapters/Chapter5/results}