### 2013-12-17 Chapters/Chapter3.tex
WIP: Finalizing chapter 3

% !TEX root = ../main.tex
% Chapter 3

\chapter{Change detection by Support Vector Machines}

\label{Chapter3} % For referencing the chapter elsewhere, use~\ref{Chapter3}

\lhead{Chapter 3. \emph{Change detection by Support Vector Machines}} % This is for the header on each page - perhaps a shortened title

This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as introduced in \Cref{Chapter4}.
The first section formulates the problem of change detection and relates it to outlier and novelty detection.
It transforms the traditional problem of outlier detection into change detection for times series data.
In \Cref{sec:one_class_classification} the problem of \acrlong{occ} is explained and discusses a number of implementations.
In the final sections of this chapter, \Cref{sec:one_class_svm}, two \gls{svm}-based \gls{occ} methods, \gls{nu-svm} and \gls{svdd}, and the influence of model parameters are further discussed in detail.

\input{./Chapters/Chapter3/change_detection_time_series}
\input{./Chapters/Chapter3/one_class_classification}
\input{./Chapters/Chapter3/one_class_svm}