### 2013-12-18 Chapters/Chapter4.tex
WIP: chapter 4

% !TEX root = ../main.tex
% Chapter 4

\chapter{Application to Human Activity Time Series Data}

\label{Chapter4} % For referencing the chapter elsewhere, use~\ref{Chapter4}

\lhead{Chapter 4. \emph{Application to Human Activity Time Series Data}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------

Ideen naam methode:
\begin{itemize}
  \item Human Activity Support Vector Change Detection
  \item Inertial Sensor Time Series Svm Temporal Segmentation
  \item Support Vector bases Human Activity Temporal Segmentation
  \item Human Activity Temporal Segmentation by Support Vector Machines: HATS-SVM
  \item HATS with One-Class SVM: HATS-OCS, HATS-OS, HATSOS
  \item OCS-HATS
  \item HATS-SVDD
\end{itemize}

\section{Experiment notes}
\begin{itemize}
  \item ICSS/CUSUM is goed in het vinden van variance changes. Niet goed in mean veranderingen.
  \item geprobeerd direct model-reset te doen na change point, maar dan kan je achteraf niet meer verder analyseren.
\end{itemize}

\begin{description}
  \item[Data gathering] Explain data gathering methods. Refer to chapter 5 and 6 for articifial and real-world details.
  \item[Model construction] Explain SVDD model construction and updating.
  \item[Model properties] Explain the SVDD properties that are used to calculate the change indication.
  \item[Change detection] Explain the interpretation of the properties, the possible (modulair) methods to give change indication
\end{description}

\TODO{Write chapter introduction/outline. Give overview of method approach and refer to each section for details of that stage.}

\TODO{Puts emphasis on two-stage framework by Takeuchi and Yamanishi \cite{takeuchi2006unifying}.}.

\TODO{Put somewhere difference and emphasis of online vs. offline/batch/post-processing}

\begin{figure}
  \centering
    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/method_setup_short.pdf}
  \caption[Method setup]{Schematic overview of the change detection method. The first step is the data gathering, described in Section \ref{sec:method_data_gathering}. Ater the pre-processing, the data is used to construct a \gls{svdd} model, as described in Section \ref{sec:method_model_construction}. Section \ref{sec:method_model_features} describes which features of the model are used for the final change indication algorithm, discussed in Section \ref{sec:method_change_detection}.}
  \label{fig:method_overview_short}
\end{figure}

\input{./Chapters/Chapter4/data_gathering}
\input{./Chapters/Chapter4/model_construction}
\input{./Chapters/Chapter4/model_features}
\input{./Chapters/Chapter4/change_detection}

 % --- OLD ---
% \input{./Chapters/Chapter4/change_indication}
% \input{./Chapters/Chapter4/overview}