### 2013-12-19 Chapters/Chapter4.tex
WIP: Chapter 4

% !TEX root = ../main.tex
% Chapter 4

\chapter{Application to Human Activity Time Series Data}

\label{Chapter4} % For referencing the chapter elsewhere, use~\ref{Chapter4}

\lhead{Chapter 4. \emph{Application to Human Activity Time Series Data}} % This is for the header on each page - perhaps a shortened title

Ideen naam methode:
\begin{itemize}
  \item Human Activity Support Vector Change Detection
  \item Inertial Sensor Time Series Svm Temporal Segmentation
  \item Support Vector bases Human Activity Temporal Segmentation
  \item Human Activity Temporal Segmentation by Support Vector Machines: HATS-SVM
  \item HATS with One-Class SVM: HATS-OCS, HATS-OS, HATSOS
  \item OCS-HATS
  \item HATS-SVDD
\end{itemize}

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{./Figures/chapter4/method_setup.pdf}
  \caption[Method setup]{Schematic overview of the change detection method. The first step is the data gathering, described in Section \ref{sec:method_data_gathering}. Ater the pre-processing, the data is used to construct a \gls{svdd} model, as described in Section \ref{sec:method_model_construction}. Section \ref{sec:method_model_features} describes which features of the model are used for the final change indication algorithm, discussed in Section \ref{sec:method_change_detection}.}
  \label{fig:method_overview}
\end{figure}

This chapter will introduce our method and setup for the temporal segmentation of human activities, using an \gls{oc-svm} approach.
The setup follows the structure as illustrated in \Cref{fig:method_overview} and further discussed in \Cref{sec:method_overview}.
It starts with the data collection phase, in which we both use artificial and real-world data.
The data sets are discussed in \Cref{sec:data_gathering}.
After (optional) pre-processing, the \gls{svdd} algorithm is used to create a model representation, which is iteratively updated.
In \Cref{sec:method_model_construction} this construction phase is discussed and shows the windowed approach.
From this constructed model we can extract some features.
In our case, we use the approximated radius $R$ of the hypersphere to detect changes.
\Cref{sec:method_model_features} goes into detail about the rationale of using that feature.
Finally, in \Cref{sec:change_detection_methods} we discuss two different, but simple, methods to detect changes in the time series data, based on the extracted radius $R$.

\input{./Chapters/Chapter4/method_overview}
\input{./Chapters/Chapter4/data_gathering}
\input{./Chapters/Chapter4/model_construction}
\input{./Chapters/Chapter4/model_features}
\input{./Chapters/Chapter4/change_detection}