### 2013-12-19 Chapters/Chapter5.tex
FrontmatteR

% !TEX root = ../main.tex
% Chapter 5

\chapter{Artificial data results}

\label{Chapter5} % For referencing the chapter elsewhere, use~\ref{Chapter5}

\lhead{Chapter 5. \emph{Artificial data results}} % This is for the header on each page - perhaps a shortened title
\begin{itemize}
  \item Problemen met SVM methode: ook na de verandering blijft de threshold nog stijgen;
  \item Veranderingen in mean geven ``extra'' change point wanneer alle data weer in het window zit. Zie \Cref{fig:camci_fixed_increasing_mean} op de 50-tallen.
\end{itemize}

In this chapter we will apply the proposed algorithm as stated in the previous chapter to artificial data sets.
For comparison, we use some of the data sets that are used by Camci~\cite{camci2010change} and Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
These data sets are generated by Gaussian noise distributions.
In \Cref{sec:artificial_data_generation} the precise generation method and characteristics of the data sets are discussed.
For our final change detection mechanism we use the \gls{rt}-method, also applied by Camci \cite{camci2010change}.
This is a simple and fast method and allows us to mainly focus on the \gls{oc-svm} model properties that should reflect changes in the data.
The results of these two methods are discussed in \Cref{sec:artificial_data_results}.

\input{./Chapters/Chapter5/artificial_data_generation}

\TODO{Add short section on quality metrics}

\input{./Chapters/Chapter5/results}