### 2014-01-06 Chapters/Chapter6.tex

% !TEX root = ../main.tex
% Chapter 6

\chapter{Real-world results}

\label{Chapter6} % For referencing the chapter elsewhere, use~\ref{Chapter6}

\lhead{Chapter 6. \emph{Real-world results}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}

This chapter discusses the real-world data sets used for this research.
In \Cref{sec:data_sets} we discuss earlier used data sets and argue that these are inappropriate for our research.
It is followed by an overview of the setup we have created to record continuous data sets from an uncontrolled environment, both in- and outdoor.
The activities performed are discussed in \Cref{sec:data_gathering}.
As discussed in \Cref{Chapter4}, the algorithm is subject to user parameters.
% In \Cref{sec:real_world_parameters} the effect and ranges of the parameters on the real world data sets are discussed.
Finally, the results of the algorithm is discussed in \Cref{sec:real_world_performance}.

\TODO{Check for final contents and ordering.}

\input{./Chapters/Chapter6/data_sets}
\input{./Chapters/Chapter6/data_gathering}
% \input{./Chapters/Chapter6/parameters}

% \TODO{Add short section on quality metrics}

\input{./Chapters/Chapter6/results}