### 2014-01-10 Chapters/Chapter2.tex
Finalizing Chapter 2, literatur review...

% !TEX root = ../main.tex
% Chapter 2

\chapter{Literature review}

\label{Chapter2} % For referencing the chapter elsewhere, use~\ref{Chapter2}

\lhead{Chapter 2. \emph{Literature review}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
% \section{Outline}
% \emph{Not intented for the reader.}
% \begin{itemize}
%   \item Literature review about Temporal Segmentation (previous draft was more about classification)
%   \item Consider methods for the context of filter-methods for classification
%   \item Take a loot at 3-4 different kind of methods for change detection:
%     \begin{itemize}
%       \item Introduction with a lot of techniques
%       \item Explain why look at a few
%       \item CUSUM - or other more traditional methods
%       \item Density-ratio estimation
%       \item Support Vector Machines (?) - if there are more sources about this
%       \item (Dimensionality reduction) --> probably not
%       \item \emph{Not to much attention to all techniques, focus is on SVM}
%     \end{itemize}
%   \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
% \end{itemize}

% \section{Statistical framework}\label{statistical-framework}

% -- Nieuwe indeling Chapter 2 --

% \begin{description}
%   \item[Intro/outline] Introduction and overview of this chapter. Put in the context of temporal classification.
%   \item[Temporal segmentation] Give overview of used temporal segmentation methods, move towards a split in model building/dimensionality reduction and change detection methods
%   \item[Dimensionality reduction] Explain that \gls{dr} can be useful in data processing. Name a few techniques, make bridge to \gls{oc-svm}.
%   \item[Change detection] Describe methods for \emph{direct} change detection. End with notion
%     \begin{description}
%       \item[Model fitting]
%       \item[Cusum]
%       \item[Density ratio]
%     \end{description}
%   \item[Support Vector Machine based methods] test

% \end{description}
As described in \Cref{Chapter1}, the focus of this research is the application of temporal segmentation to accelerometer data.
This is with the assumption that explicit temporal segmentation can aid the process of temporal classification.

\TODO{Give examples of temporal classification techniques}

In recent years a lot is research is performed in the classification of human activities, recorded by on-body sensors.
Especially since smartphones with accelerometer sensors became more available, research has focused on recordings from these devices.

\TODO{Include highlights from old-classification text}

In the following of this chapter we will start with an overview of statistical models used in this field of study.
That section will act as a starting point to look at the different types, applications and methods of temporal segmentation.
The next section, \Cref{sec:literature_review_temporal_segmentation} follows the distinction made and looks at a collection of temporal segmentation and change detection methods.
It will also relate the concept of \gls{occ} to temporal segmentation and change detection.
The final section of this chapter, \Cref{sec:literature_review_svm}, will discuss a specific class of algorithms for \gls{occ}, namely the methods that use \glspl{svm} for the classification task.

\input{./Chapters/Chapter2/statistical_models.tex}
\input{./Chapters/Chapter2/temporal_segmentation.tex}
% \input{./Chapters/Chapter2/change_detection_methods.tex}

% -- OLD --

%----------------------------------------------------------------------------------------
% CUSUM and GLR
% \input{./Chapters/Chapter2/cusum}

%----------------------------------------------------------------------------------------
% Density Ratio Estimation
\input{./Chapters/Chapter2/density_ratio}

%----------------------------------------------------------------------------------------
% Support Vector Machines
\input{./Chapters/Chapter2/svm}