### 2014-01-15 Chapters/Chapter3/one_class_svm.tex
Finished feedback Chapter 3

% !TEX root = ../../main.tex
\section{One-Class Support Vector Machine}\label{sec:one_class_svm}
In this section we will discuss the details of an \gls{svm} and \gls{oc-svm} implementations.
The classical implementation of an \gls{svm} is to classify a dataset in two distinct classes.
This is a common use case, although sometimes there is no training data for both classes available.
Still, one would like to classify new data points as regular, in-class, or out-of-class, \eg in the case of a novelty detection.
With that problem only data examples from one class are available and the objective is to recognize new data points that are not part of that class.
This unsupervised learning problem is closely related to density estimation.
In that context, the problem can be the following.
Assume an underlying probability distribution $P$ and a data point $\vectorsym{x}$ drawn from this distribution.
The goal is to find a subset $S$ of the input space, such that the probability that $\vectorsym{x}$ lies inside of $S$ equals some predetermined value $\nu$ between $0$ and $1$ \cite{scholkopf1999support}.

In the following of this section we will start with a discussion of traditional \glspl{svm}.
In \Cref{subsec:kernels} we will show how kernels allow for non-linear decision functions.
That is followed by two different implementations of \gls{oc-svm}: \gls{nu-svm} in \Cref{subsec:nu-svm} by Sch\"olkopf \etal \cite{scholkopf1999support}, which closely follows the above problem statement regarding density estimation, and \gls{svdd} by Tax and Duin \cite{tax1999support} in \Cref{subsec:oc-svm-svdd}.
The final part of this section discusses the influence of model parameters on the performance of \glspl{svm}

%--------------------------------------------
\subsection{Support Vector Machine}\label{subsec:svm}
We will first discuss the traditional two-class \gls{svm} before we consider the one-class variant, as introduced by Cortes and Vapnik in \cite{cortes1995support}.
Consider a labeled data set $\Omega = \{ (x_1, y_1),\allowbreak (x_2, y_2), \dots , (x_n, y_n) \}$; points $x_i \in \mathbb{I}$ in a (for instance two-dimensional) space where $x_i$ is the $i$-th input data point and $y_i \in \{-1, 1\}$ is the $i$-th output value, indicating the class membership.

An \gls{svm} can create a boundary between linear-separable data points, making it a binary linear classifier.
More flexible non-linear boundaries can be obtained by the use of a non-linear function $\phi(x)$, as illustrated in \Cref{fig:kernel_mapping}.
This function maps the input data from space $\mathcal{I}$ to a higher dimensional space $\mathcal{F}$.
The \gls{svm} can create a linear separating hyperplane in the space $\mathcal{F}$ that separates the data points from the two classes.
When the hyperplane is projected on to the (lower) original input space $\mathcal{I}$ it creates a non-linear separating curve.
A schematic overview of this process is illustrated in \Cref{fig:svm_mapping_spaces}.
The original data points $x \in \mathcal{I}$ are projected by $\phi(x)$ to the points $z \in \mathcal{F}$.
In that space the linear hyperplane is constructed by weights $w$.
This results in a non-linear hyperplane $y$ in input space $\mathcal{I}$.
The mapping and projection of data points can be efficient (and implicit) performed by using the kernel trick, which is discussed in \Cref{subsec:kernels}.

\begin{figure}
\centering
  \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_kernel_mapping.png}
  \caption[Kernel mapping]{Visualization of an \gls{svm} in 2D.
  The non-linear boundary in the input space $\mathcal{I}$ (left) is transformed to a linear boundary in the feature space $\mathcal{F}$ (right) by mapping the data points with the function $\phi$.
  The kernel trick uses a function $K$ which performs an implicit mapping. Image from Wikipedia.org}
  \label{fig:kernel_mapping}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_mapping_spaces.pdf}
  \caption[Mapping spaces in \gls{svm}]{Schematic visualization of the non-linear mapping of \glspl{svm}.
  The input vector $\vectorsym{x}$ from input space $\mathcal{I}$ is mapped by a non-linear function $\phi(\vectorsym{x})$ to a feature vector $\vectorsym{z}$ in the high dimensional feature space $\mathcal{F}$.
  The weights of $\vectorsym{w}$ create a linear separating hyperplane, which maps the high dimensional vector to the predicted outcome $\vectorsym{y}$.}
  \label{fig:svm_mapping_spaces}
\end{figure}

The separating hyperplane is represented by
\begin{equation}
w^T x + b = 0,
\end{equation}
with $w \in \mathcal{F}$ and $b \in \mathbb{R}$.
The hyperplane that is created determines the \emph{margin} between the classes; the minimal distance from the data points to the hyperplane.
In geometric sense, $w$ is the normal vector indicating the direction of the hyperplane and $\frac{b}{\lVert{w}\rVert}$ determines the offset of the hyperplane from the origin.
Since the size of the margin is equal to $\frac{2}{\lVert{w}\rVert}$, the maximum-margin hyperplane is found by minimizing $\lVert{w}\rVert$.
The data points which lie on the boundary of the margin are the \emph{support vectors}.
This geometrical interpretation is illustrated in \Cref{fig:svm_hyperplane}.
All data points for which $y_i = -1$ are on one side of the hyperplane and all other data points (for which $y_i = 1$) are on the other side.
The minimal distance from a data point to the hyperplane is equal for both classes.
Minimizing $\lVert{w}\rVert$ results in a \emph{maximal margin} between the two classes.
Thus, the \gls{svm} searches for a maximal separating hyperplane.

\begin{figure}
\centering
  \includegraphics[width=0.5\textwidth]{./Figures/chapter3/svm_separating_plane_with_margin.png}
  \caption[\gls{svm} and the separating hyperplane]{Illustration of the separating hyperplane of an \gls{svm}.
  Here $w$ is the normal vector for the separating hyperplane and size of the margin is $\frac{2}{\lVert{w}\rVert}$.
  Image from Wikipedia.org}
  \label{fig:svm_hyperplane}
\end{figure}

With every classification method there is a risk of overfitting.
In that case the random error or noise of the data set is described instead of the underlying data.
The \gls{svm} classifier can use a \emph{soft margin} by allowing some data points to lie within the margin, instead of on the margin or farther away from the hyperplane.
For this it introduces \emph{slack variables} $\xi_i$ for each data point and the constant $C > 0$ determines the trade-off between maximizing the margin and the number of data points within that margin (and thus the training errors).
The slack variables are a generalization to minimize the \emph{sum of deviations}, rather than the \emph{number} of incorrect data points \cite{cherkassky2007learning}\footnote{if $\xi_i$ is chosen to be an indicator function, it still minimizes the \emph{number} of incorrect data points.}.
The objective function for an \gls{svm} is the following minimization function:

\begin{equation}\label{eq:svm_objective}
  \operatorname*{min}_{w,\ b,\ \xi_i} \frac{ \lVert{w}\rVert^2 }{2} + C \sum_{i=1}^n \xi_i
\end{equation}
\begin{equation}
  \begin{multlined}
  \mbox{ subject to: } \\*
  \begin{aligned}
  y_i( w^T \phi(x_i) + b) \geq \: & 1 - \xi_i & \mbox{ for all } i = 1, \dots, n \\*
   & \xi_i \geq 0 & \mbox{ for all } i = 1, \dots, n\\*
  \end{aligned}
  \end{multlined}
\end{equation}

This minimization problem can be solved using \gls{qp} and scales in the number of dimensions $d$, since $w \in \mathbb{R}^d$.
In its dual formulation, as shown in \Cref{eq:svm_lagrange}, the non-negative Lagrange multipliers $\alpha_i$ are introduced.
Using this formulation the problem scales in the number of data points $n$, since $\vectorsym{\alpha} \in \mathbb{R}^n$.
% In the dual formulation the problem scales with the number of training examples $n$ instead of the dimensionality $d$ of the samples.
Solving this problem directly in the high dimensional feature space $\mathcal{F}$ makes it intractable.
% The linear approximation function corresponds to the kernel function in the dual formulation.
% Solving this dual formulation is equivalent to solving the primal formulation \cite{cherkassky2007learning}.
It is shown that solving the dual formulation is equivalent to solving the primal form \cite{cherkassky2007learning}.
% In the dual formulation the non-negative Lagrange multipliers $a_i$ are introduced and the decision function becomes:
The linear classifier decision function in the dual form is optimized over $\alpha_i$:
\begin{equation}\label{eq:svm_lagrange}
  f(x) = \operatorname{sgn}( \sum_{i=1}^n \left\{ \alpha_i y_i K(x, x_i) + b \right\} ),
\end{equation}
where $K(x, x_i) = \phi(x)^T\phi(x_i)$ is the dot product of data objects in feature space $\mathcal{F}$ (which is further discussed in \Cref{subsec:kernels}).
Here every data point in $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
Since it is shown that under certain circumstances \glspl{svm} resembles sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.

Using this formulation two important properties arise \cite{flach2012machine}:
\begin{enumerate}
  \item Searching for the maximum margin decision boundary is equivalent to searching for the support vectors; they are the training examples with non-zero Lagrange multipliers.
  \item The optimization problem is entirely defined by pairwise dot products between training examples: the entries of the kernel matrix $K$.
\end{enumerate}
An effect of the first property, combined with the equality to sparse representations, is that \glspl{svm} often have good results, even in the case of high dimensional data or limited training examples \cite{cherkassky2007learning}.
The second property is what enables an powerful adaptation of \glspl{svm} to learn non-linear decision boundaries.
The workings of the kernel matrix $K$ and the non-linear boundaries are discussed in the following section.

%--------------------------------------------

\subsection{Kernels}\label{subsec:kernels}
In the previous section, the mapping function $\phi(x)$ and the kernel function $K$ were briefly mentioned.
The decision function in \Cref{eq:svm_lagrange} only relies on the dot products of mapped data points in the feature space $\mathcal{F}$ (\ie all pairwise distances between the data points in that space).
It shows \cite{flach2012machine} that for any function $K$ that corresponds to a dot product in feature space $\mathcal{F}$, without an explicit mapping to the higher dimension $\mathcal{F}$, the dot products can be substituted by that kernel function $K$.
This is introduced by Aizerman~\etal \cite{aizerman1964theoretical} and applied to \glspl{svc} by Vapnik~\cite{vapnik1998statistical}, known as the `kernel trick':
\begin{equation}
  K(\vectorsym{x}, \vectorsym{x}') = (\vectorsym{z} \cdot \vectorsym{z}') = \phi(\vectorsym{x}) \cdot \phi(\vectorsym{x}')
\end{equation}
where vectors $\vectorsym{z}$ and $\vectorsym{z}'$ are projections of data objects $\vectorsym{x}$ and $\vectorsym{x}'$ through $\phi(\vectorsym{x})$ on the features space $\mathcal{F}$.
% The dot product $\phi(\vectorsym{x}) \cdot \phi(\vectorsym{x}')$ can be replaced by $K(\vectorsym{x}, \vectorsym{x}')$, which is known as the `kernel trick' and found by Vapnik~\cite{vapnik1998statistical}.
% \begin{equation}
  % K(\vectorsym{x}, \vectorsym{x}') = \sum_{j=1}^n g_j(\vectorsym{x}) g_j(\vectorsym{x}')
  % K(\vectorsym{x}, \vectorsym{x}') = \sum_{j=1}^n k(\vectorsym{x}, \vectorsym{x}_j)
% \end{equation}
% of $n$ basis functions $k(\vectorsym{x}, \vectorsym{x}')$.
Note that the evaluation of dot products in the feature space $\mathcal{F}$ between vectors is performed indirectly via the evaluation of the kernel $K$ between (support) vectors in the input space $\mathcal{I}$.
% This is known as the \emph{kernel trick} and gives the \gls{svm} the ability to create non-linear decision function without high computational complexity.
This is gives the \gls{svm} the ability to create non-linear decision function without high computational complexity.

The kernel function $K$ can have different forms, such as linear, polynomial and sigmoidal but the most used (and flexible) form is the Gaussian \gls{rbf}.
The inner product kernel with \gls{rbf} basis functions have the form
\begin{equation}
  % k(x, x') = \operatorname*{sign} \left(  \sum_{i=1}^{n} \alpha_i \operatorname*{exp} \left\{ \frac{\rVert \vectorsym{x}_j - \vectorsym{x} \rVert ^2}{\sigma^2} \right\} \right),
  K(\vectorsym{x}, \vectorsym{x}') = \operatorname*{exp} \left\{ - \frac{\rVert \vectorsym{x} - \vectorsym{x}' \rVert ^2}{2 \sigma^2} \right\},
\end{equation}
where $\sigma$ defines the Gaussian width and $\rVert \vectorsym{x} - \vectorsym{x}' \rVert ^2$ is the dissimilarity in Euclidean distance.
% The inner product kernel $K$ then becomes
% \begin{equation}
  % K(\vectorsym{x}, \vectorsym{x}') = \operatorname*{exp} \left\{ - \frac{ \rVert \vectorsym{x} - \vectorsym{x}' \rVert ^2}{\sigma^2} \right\}.
% \end{equation}

The kernel $K$ maps input space $\mathcal{I}$ to the feature space $\mathcal{F}$ which is a \gls{rkhs} of (theoretically) infinite dimensions.
As Smola \etal \cite{smola1998connection} state, this Gaussian kernel yields good performance, especially when no assumptions can be made about the data.
As an explanation, they show a correspondence between learning \glspl{svm} with \gls{rbf} kernels and regularization operators.
This may give insights in why \glspl{svm} have been found to exhibit high generalization ability (by learning with few training objects).

% The number of basis functions, the center parameters that correspond with support vectors, and the weights in the output layer are all automatically determined via the optimal hyperplane in features space $\mathcal{F}$ \cite{cherkassky2007learning}.
% The center parameters that correspond with support vectors and the weights in the output layer are automatically determined via the optimal hyperplane in features space $\mathcal{F}$ \cite{cherkassky2007learning}.
The width parameter $\sigma$ is equal for all kernel functions and is set a priori and determines the flexibility and complexity of the boundary.
In \Cref{subsec:svm_model_parameters} this (hyper)parameter for an \gls{svm} is further discussed.

The mapping from input space $\mathcal{I}$ to $\mathcal{F}$ via $\phi(x)$ is subject to some continuity assumptions.
This general assumption in pattern recognition, states that two near objects in feature space should also resemble each other in ``real life'' \cite{tax2001one}.
Thus, objects which are close in feature space should be close in the original input space.
When this assumption does not hold, the example objects would be scatter through the feature space and finding a decision function becomes very complex.

% -- Quotes and refs to use --x

% sparsity, \gls{rkhs}: \cite{girosi1998equivalence} \\

% Use Chapter 2 from ``Learning with kernels'', \cite{scholkopf2002learning}.

%--------------------------------------------

\subsection{\acrlong{nu-svm}}\label{subsec:nu-svm}
The first of the \gls{oc-svm} methods we will discuss is often referred to as \gls{nu-svm} and introduced by Sch\"olkopf \etal \cite{scholkopf1999support}.
Instead of estimating the density function of an distribution $P$, it focuses on an easier problem: the algorithm finds regions in the input where the ``probability density lives''.
This results in a function such that most of the data is in the region where the function is nonzero.

The constructed decision function $f(x)$ resembles the function discussed in \Cref{subsec:svm}.
It returns the value $+1$ in a (possibly small) region capturing most of the data points, and $-1$ elsewhere.
The method maps the data points from input space $\mathcal{I}$ to a feature space $\mathcal{F}$ (following classical \glspl{svm}).
In that space the data points are separated from the origin by a hyperplane, with maximal margin.
Since the mapping to the feature space is based on dot products of the data points, outlier objects (which are dissimilar from the training set) will be closer to the origin.
Thus, maximizing the distance from the hyperplane to the origin increases the discriminating ability of the decision function.
Furthermore, it holds an intuitive relationship with the classical two-class \gls{svm}.

For a new data points $x$, the function value $f(x)$ determines whether the data point is part of the distribution (\ie the value is $+1$) or a novelty (\ie the value is $-1$).
The hyperplane is represented by $g(x) = w \cdot \phi(x) + \rho = 0$ and the decision function is $f(x) = \operatorname{sgn}(g(x))$.
This hyperplane and the separation from the origin is illustrated in \Cref{fig:nu-svm}.

\begin{figure}
  \centering
    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/nu-svm.pdf}
  \caption[\gls{nu-svm}]{Graphical representation of \gls{nu-svm}. The separating hyperplane $w \cdot \phi(x_i) + \rho = 0$ create a maximal margin, in the feature space, between the data points and the origin. Slack variables $\xi_i$ are used to create a soft margin.}
  \label{fig:nu-svm}
\end{figure}

The objective function to find the separating hyperplane is the following minimization function, which can be solved using \gls{qp}:

\begin{equation}\label{eq:nu-svm_objective}
  \operatorname*{min}_{w,\ \xi_i,\ \rho } \frac{\lVert w \rVert ^2}{2} + \frac{1}{\nu n} \sum_{i=1}^n \xi_i - \rho
\end{equation}
\begin{equation}
  \begin{multlined}
    \mbox{ subject to: } \\
    \begin{aligned}
      (w \cdot \phi(x_i)) \geq \: & \rho - \xi_i & \mbox{ for all } i = 1, \dots, n \\
      & \xi_i \geq 0 & \mbox{ for all } i = 1, \dots, n \\
    \end{aligned}
  \end{multlined}
\end{equation}

The decision function in the dual formulation with Lagrange multipliers is denoted as:
\begin{equation}\label{eq:nu-svm_lagrange}
f(x) = \operatorname{sgn}((w \cdot \phi(x_i)) - \rho) = \operatorname{sgn}( \sum_{i=1}^n \alpha_i K(x, x_i) - \rho)
\end{equation}

In the classical \gls{svm} objective function, as denoted in \Cref{eq:svm_objective}, the parameter $C$ decided the smoothness of the boundary, with respect to the slack variables $\xi_i$.
In the formulation of \gls{nu-svm} the equivalent parameter is $\nu \in (0,1)$ (hence the name).
It characterizes the solution in two ways:
\begin{enumerate}
  \item $\nu$ is an upper bound on the fraction of outliers, \ie training examples regarded as out-of-class.
  \item $\nu$ is a lower bound on the fraction of \glspl{sv}, \ie training examples with a nonzero Lagrange multiplier $\alpha_i$.
\end{enumerate}
When $\nu$ approaches $0$, the penalty factor for nonzero Lagrange multipliers ($\frac{1}{\nu n}$) becomes infinite, and thus the solution resembles a \emph{hard margin} boundary.

This method creates a \emph{hyperplane}, characterized by $w$ and $\rho$, that separates the data with maximal margin from the origin in the feature space $\mathcal{F}$.
In the following section we will discuss an alternative method, which uses an circumscribing \emph{hypersphere} to characterize the training data.
The region inside the hypersphere indicates the region $S$ where the probability that a data point drawn from $P$ is equal to $\nu$.

%--------------------------------------------

\subsection{\acrlong{svdd}}\label{subsec:oc-svm-svdd}
The method introduced by Tax and Duin \cite{tax1999support}, known as \acrlong{svdd}, follows a spherical instead of planar approach.
The boundary, created in feature space $\mathcal{F}$, forms a hypersphere around the (high density region of the) data.
The volume of this hypersphere is minimized to get the smallest enclosing boundary.
The chance of accepting outlier objects is thereby also minimized \cite{tax2003online}.
By allowing outliers using slacks variables, in the same manner as classical \gls{svm} and \gls{nu-svm}, a soft margin is constructed.

\begin{figure}
  \centering
    \includegraphics[width=0.8\textwidth,keepaspectratio]{./Figures/chapter3/nu-vs-svdd.pdf}
  \caption[Difference \gls{nu-svm} and \gls{svdd}]{Graphical representation of the difference between \gls{nu-svm} (left) and \gls{svdd} (right). Note that for the sake of simplicity the kernel functions are not applied.}
  \label{fig:nu-vs-svdd}
\end{figure}

The constructed hypersphere is characterized by the center $\mathbf{a}$ and radius $R > 0$ as distance from the center to (any data point that is a \gls{sv} on) the boundary, for which the volume, and thus the radius $R$, will be minimized.
The center $\mathbf{a}$ is a linear combination of the support vectors.
Like the classical \gls{svm} and \gls{svdd} it can be required that all the distances from the data points $x_i$ to the center $\mathbf{a}$ are strict less than $R$ (or equivalent measure).
A soft margin can be allowed by using slack variables $\xi_i$.
In that case, the penalty is determined by $C$ and the minimization is expressed as \Cref{eq:svdd_objective}.
This principle is illustrated in the right image or \Cref{fig:nu-vs-svdd}.
Instead of a separating hyperplane, constructed by \gls{nu-svm} and illustrated on the left of the Figure, the \gls{svdd} creates a hypersphere (in the illustration a circle) around the data points.
By using kernel functions (\eg the \gls{rbf}) the hyperspheres in the high dimensional feature space $\mathcal{F}$ corresponds to a flexible and tight enclosing boundary in input space $\mathcal{I}$.
Possible resulting closed boundaries are illustrated in \Cref{fig:svdd-boundary}.
This enclosing boundary is obtained by minimizing the following error function $L$ which contains the volume of the hypersphere and the distance from the boundary to the outlier objects:
\begin{equation}\label{eq:svdd_objective}
  L(R, \mathbf{a}, \vectorsym{\xi}) = R^2 + C \sum_{i=1}^n \xi_i
\end{equation}
\begin{equation}
  \begin{multlined}
    \mbox{ subject to: } \\
    \begin{aligned}
      \lVert x_i - \mathbf{a} \rVert ^ 2 \leq \: & R^2 + \xi_i & \mbox{ for all } i = 1, \dots, n \\
      & \xi_i \geq 0 & \mbox{ for all } i = 1, \dots, n \\
    \end{aligned}
  \end{multlined}
\end{equation}
In the dual Lagrangian formulation of this error function $L$ the multipliers $\vectorsym{\alpha}$ are maximized:
\begin{equation}\label{eq:svdd_lagrange}
  L = \sum_{i} \alpha_i(x_i \cdot x_i) - \sum_{i,j} \alpha_i \alpha_j(x_i \cdot x_j)
\end{equation}
\begin{equation}
  \begin{multlined}
    \mbox{ subject to: } \\
    \begin{aligned}
    0 \le \alpha_i \le C, \: \sum_{i} \alpha_i = 1
    \end{aligned}
  \end{multlined}
\end{equation}

In the maximization of \Cref{eq:svdd_lagrange} a large fraction of the multipliers $\alpha_i$ become zero.
The small fraction for which $\alpha_i > 0$ are called the \glspl{sv} and these objects lie on the boundary of the description.
The center of the hypersphere only depends on this small number of \glspl{sv} and the objects for which $\alpha_i = 0$ can be discarded from the solution.
Testing the membership of a (new) object $\vectorsym{z}$ is done by determining if the distance to the center $\mathbf{a}$ of the sphere is equal or smaller to the radius $R$:
\begin{equation}\label{eq:svdd_test_object}
  \lVert \vectorsym{z} - \mathbf{a} \rVert ^2 = (\vectorsym{z} \cdot \vectorsym{z}) - 2 \sum_{i} \alpha_i(\vectorsym{z} \cdot \vectorsym{x}_i) + \sum_{i,j} \alpha_i \alpha_j (\vectorsym{x}_i \cdot \vectorsym{x}_j) \le R^2
\end{equation}
As with \Cref{eq:svm_lagrange}, the solution of this equation only relies on dot products between the data points in $\vectorsym{x}$ and $\vectorsym{z}$.
This means that the kernel projection and trick, as discussed in \Cref{subsec:kernels}, can be applied to \gls{svdd} as well \cite{tax1999support,tax2002uniform}.

Because the Gaussian \gls{rbf} often yields good (\ie tight) boundaries, this set of kernels functions is commonly used:
\begin{equation}
  (\vectorsym{x} \cdot \vectorsym{y}) \rightarrow K(\vectorsym{x},\vectorsym{y}) = \operatorname{exp}\left(- \frac{\lVert \vectorsym{x} - \vectorsym{y} \rVert ^2}{\sigma^2}\right)
\end{equation}
Using this kernel function, the Lagrangian error function $L$ of \Cref{eq:svdd_lagrange} changes to:
\begin{equation}\label{eq:svdd_lagrange_kernel}
  L = 1 - \sum_{i} \alpha_i^2 - \sum_{i \ne j} \alpha_i \alpha_j K(x_i, x_j)
\end{equation}
Using \Cref{eq:svdd_test_object}, the following kernel formulation needs to hold for a new object $\vectorsym{z}$ to lie within the hypersphere:
\begin{equation}\label{eq:svdd_inequality}
  \sum_{i} \alpha_i K(\vectorsym{z}, x_i) \le \frac{1}{2} \left( 1 - R + \sum_{i,j} \alpha_i \alpha_j K(x_i, x_j) \right)
\end{equation}
When the Gaussian \gls{rbf} kernel is applied, and in case the data is preprocessed to have unit length (for the \gls{nu-svm} solution), the two different \gls{oc-svm} implementations \gls{nu-svm} and \gls{svdd} are shown to have identical solutions \cite{tax2002uniform,scholkopf2002learning}

\begin{figure}
  \centering
    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/svdd-boundary.pdf}
  \caption[\gls{svdd} boundary]{The \gls{svdd} method trained on a banana-shaped data set with different sigma-values for the \gls{rbf} kernel. Solid circles are support vectors, the dashed line is the boundary. Image by Tax \cite{tax2001one}.}
  \label{fig:svdd-boundary}
\end{figure}

%--------------------------------------------

\subsection{SVM model parameters}\label{subsec:svm_model_parameters}
\gls{svm}-model selecting and tuning depends on two type of parameters \cite{cherkassky2007learning}:
\begin{enumerate}
  \item Parameters controlling the `margin' size,
  \item Model parameterization, \eg the kernel type and complexity parameters.
  For the \gls{rbf} kernel the width parameter determines the model complexity.
\end{enumerate}

In case of a \gls{rbf} kernel, the width parameter $\sigma$ determines the flexibility and complexity of the boundary.
The value of this parameter greatly determines the outcomes of the algorithm (\eg \gls{svdd}) as illustrated in \Cref{fig:svdd-boundary}.
With a small value for the kernel width $\sigma$, each data point will tend to be used as a support vector (for almost all $\alpha_i > 0$) and the \gls{svdd} solution resembles a Parzen density estimation.
For large values of $\sigma$, the solution will resemble the original hypersphere solution (in contrast with a tight boundary around the data).
With a large value for the width $\sigma$, the boundary approximates the spherical boundary.
The influence of the $\sigma$ parameter on the \gls{svdd} solution is illustrated in \Cref{fig:svdd-boundary-sigma}.

\begin{figure}
  \centering
    \includegraphics[width=0.8\textwidth,keepaspectratio]{./Figures/chapter3/svdd-parameter-sigma.pdf}
  \caption[\gls{svdd} boundary]{The \gls{svdd} method trained on a banana-shaped data set with different $\sigma$-values for the \gls{rbf} kernel. Solid circles are support vectors, the dashed line is the boundary.}
  \label{fig:svdd-boundary-sigma}
\end{figure}

As discussed in \Cref{subsec:nu-svm}, the \gls{svm} parameter $C$ (or $\nu$ in case of \gls{nu-svm}) is of high influence on the ``smoothness'' of the decision function.
It acts as an upper bound to the fraction of outliers and as a lower bound to the fraction of \glspl{sv}.
A more detailed discussion of the influence of the \gls{svm} model parameters can be found in Section~9.8 of \cite{cherkassky2007learning} and Section~7.4 from \cite{flach2012machine}.
A detailed discussion of the $\nu$ and kernel parameters can be found in \cite{scholkopf2002learning}.

The following chapter will discuss our proposed method, which incorporates the \gls{svdd} algorithm.
It relates the \gls{oc-svm} model construction to outlier detection and eventually change detection, leading to finding a temporal segmentation of time series data.

% -- Notes --
% \begin{itemize}
  % \item Low target rejection rate $f_{T-}$ and low outlier acceptance rate $f_{O+}$. When only target examples are present, the first one can be estimated by the number of support vectors that we obtain as a solution of Lagrangian~\ref{eq:svdd_lagrange_kernel}.
% \end{itemize}

% -- Literature --

% ``A geometric approach to support vector machine (SVM) classification'' \cite{mavroforakis2006geometric}. 136, 2006 \\

% ``Least squares one-class support vector machine'' \cite{choi2009least}. 27, 2009 \\

% ``On simple one-class classification methods'' \cite{noumir2012simple}. 2012  --> decouples radius and center optimization, gives fast approximations instead of precise results. \\

% ``Choosing a small width of the kernels leads to high generalization error as it effectively decouples the separate basis functions of the kernel expansion into very localized functions which is equivalent to memorizing the data, whereas a wide kernel tends to oversmooth.'' \cite{smola1998connection}. \\

% ``For small values of $\sigma$ almost all $\alpha_i >0$ and the \gls{svdd} resembles a Parzen density estimation.'' \cite{tax2002uniform} page 4, \cite{tax1999support} page 4. \\