### 2014-01-15 Chapters/Chapter4.tex
Decide on OCS-HTS. Working on algorithm in chapter 4

% !TEX root = ../main.tex
% Chapter 4

\chapter{Application to Human Activity Time Series Data}

\label{Chapter4} % For referencing the chapter elsewhere, use~\ref{Chapter4}

\lhead{Chapter 4. \emph{Application to Human Activity Time Series Data}} % This is for the header on each page - perhaps a shortened title

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{./Figures/chapter4/method_setup.pdf}
  \caption[Method setup]{Schematic overview of the change detection method. The first step is the data gathering, described in Section \ref{sec:method_data_gathering}. Ater the pre-processing, the data is used to construct a \gls{svdd} model, as described in Section \ref{sec:method_model_construction}. Section \ref{sec:method_model_features} describes which features of the model are used for the final change indication algorithm, discussed in Section \ref{sec:method_change_detection}.}
  \label{fig:method_overview}
\end{figure}

This chapter will introduce our method, which we named \gls{ocs-hats}, and setup for the temporal segmentation of human activities, using an \gls{oc-svm} approach.
It shows our method to apply a \gls{oc-svm} based temporal segmentation method to inertial sensor time series data, often used in the context of \acrlong{har}.
As far as we know it is the first in its kind, especially on the application of \gls{svdd} and \gls{svcpd} to real-world \gls{har} data.
The setup follows the structure as illustrated in \Cref{fig:method_overview} and further discussed in \Cref{sec:method_overview}.
It starts with the data collection phase, in which we both use artificial and real-world data.
The data sets are discussed in \Cref{sec:data_gathering}.
After (optional) pre-processing, the \gls{svdd} algorithm is used to create a model representation, which is iteratively updated.
In \Cref{sec:method_model_construction} this construction phase is discussed and shows the windowed approach.
From this constructed model we can extract some features.
In our case, we use the approximated radius $R$ of the hypersphere to detect changes.
\Cref{sec:method_model_features} goes into detail about the rationale of using that feature.
Finally, in \Cref{sec:method_change_detection} we discuss two different, but simple, methods to detect changes in the time series data, based on the extracted radius $R$.

\input{./Chapters/Chapter4/method_overview}
\input{./Chapters/Chapter4/data_gathering}
\input{./Chapters/Chapter4/model_construction}
\input{./Chapters/Chapter4/model_features}
\input{./Chapters/Chapter4/change_detection}