### 2014-01-17 main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Thesis
% LaTeX Template
% Version 1.3 (21/12/12)
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original authors:
% Steven Gunn
% http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/
% and
% Sunil Patel
% http://www.sunilpatel.co.uk/thesis-template/
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
% Note:
% Make sure to edit document variables in the Thesis.cls file
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt, a4paper, oneside]{Thesis} % Paper size, default font size and one-sided paper

\graphicspath{{./Pictures/}} % Specifies the directory where pictures are stored

\usepackage[square, numbers, comma, sort&compress]{natbib} % Use the natbib
%reference package - read up on this to edit the reference style; if you want
%text (e.g. Smith et al., 2012) for the in-text references (instead of
%numbers), remove 'numbers'
\usepackage{isomath}
\usepackage{paralist}
\usepackage{tabulary}
\usepackage{cleveref}
\usepackage{subcaption}
\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue -
%change to black if annoying
\title{\ttitle} % Defines the thesis title - don't touch this

% Use double bar notations: http://tex.stackexchange.com/questions/43008/absolute-value-symbols/43009#43009
\usepackage{mathtools}
\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%

\usepackage{algorithm}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

% Swap the definition of \abs* and \norm*, so that \abs
% and \norm resizes the size of the brackets, and the
% starred version does not.
\makeatletter
\let\oldabs\abs
\def\abs{\@ifstar{\oldabs}{\oldabs*}}
%
\let\oldnorm\norm
\def\norm{\@ifstar{\oldnorm}{\oldnorm*}}
\makeatother
% End of double bar notations

\DeclareMathOperator*{\argminold}{\arg\!\min}
\DeclareMathOperator*{\argmaxold}{\arg\!\max}

% \usepackage{hyperref}
\usepackage[nomain,acronym,xindy,toc]{glossaries}
\renewcommand*{\glstextformat}[1]{\textcolor{black}{#1}}
\makeglossaries
\usepackage[xindy]{imakeidx}
\makeindex

\usepackage{templates/math_shorthands}

\begin{document}
\include{glossary-acronyms}
\include{glossaries}

\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages

\setstretch{1.3} % Line spacing of 1.3

% Define the page headers using the FancyHdr package and set up for one-sided printing
\fancyhead{} % Clears all page headers and footers
\rhead{\thepage} % Sets the right side header to show the page number
\lhead{} % Clears the left side page header

\pagestyle{fancy} % Finally, use the "fancy" page style to implement the FancyHdr headers

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % New command to make the lines in the title page

\newcommand{\TODO}[1]{\emph{\textbf{*** TODO: {#1} ***}}}

\ExplSyntaxOn
\newcommand\latinabbrev[1]{
  \peek_meaning:NTF . {% Same as \@ifnextchar
    #1\@}%
  { \peek_catcode:NTF a {% Check whether next char has same catcode as \'a, i.e., is a letter
      #1.\@\ }%
    {#1.\@\ }}}
\ExplSyntaxOff

%Omit final dot from each def.
\def\eg{\latinabbrev{e.g}}
\def\etal{\latinabbrev{\textit{et al}}}
\def\etc{\latinabbrev{etc}}
\def\ie{\latinabbrev{i.e}}

% PDF meta-data
\hypersetup{pdftitle={\ttitle}}
\hypersetup{pdfsubject=\subjectname}
\hypersetup{pdfauthor=\authornames}
\hypersetup{pdfkeywords=\keywordnames}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage}
\begin{center}

\textsc{\LARGE \univname}\\[1.5cm] % University name
\textsc{\Large Master's Degree Thesis}\\[0.5cm] % Thesis type

\HRule \\[0.4cm] % Horizontal line
{\huge \bfseries \ttitle}\\[0.4cm] % Thesis title
\HRule \\[1.5cm] % Horizontal line

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
\href{http://rvlasveld.github.io}{\authornames}, \href{http://www.uu.nl}{\textsc{UU}} % Author name - remove the \href bracket to remove the link
\end{flushleft}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{flushright} \large
\emph{Supervisors:} \\
\href{http://www.cs.uu.nl/staff/gv.html}{Dr. G.A.W. \textsc{Vreeswijk}, \href{http://www.uu.nl}\textsc{UU}}  \\ % Supervisor name - remove the \href bracket to remove the link
\href{http://www.annevanrossum.com}{A.C. \textsc{van Rossum}, \href{http://www.almende.com}\textsc{Almende B.V.}} % Supervisor name - remove the \href bracket to remove the link
\end{flushright}
\end{minipage}\\[3cm]

\large \textit{A thesis submitted in fulfilment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
\textit{in the}\\[0.4cm]
\facname\\\deptname\\[2cm] % Research group name and department name

{\large \today}\\[4cm] % Date
%\includegraphics{Logo} % University/department logo - uncomment to place it

\vfill
\end{center}

\end{titlepage}

%----------------------------------------------------------------------------------------
%	DECLARATION PAGE
%	Your institution may give you a different text to place here
%----------------------------------------------------------------------------------------

\Declaration{

\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics

I, \authornames, declare that this thesis titled, '\ttitle' and the work presented in it are my own. I confirm that:

\begin{itemize}
\item[\tiny{$\blacksquare$}] This work was done wholly or mainly while in candidature for a research degree at this University.
\item[\tiny{$\blacksquare$}] Where any part of this thesis has previously been submitted for a degree or any other qualification at this University or any other institution, this has been clearly stated.
\item[\tiny{$\blacksquare$}] Where I have consulted the published work of others, this is always clearly attributed.
\item[\tiny{$\blacksquare$}] Where I have quoted from the work of others, the source is always given. With the exception of such quotations, this thesis is entirely my own work.
\item[\tiny{$\blacksquare$}] I have acknowledged all main sources of help.
\item[\tiny{$\blacksquare$}] Where the thesis is based on work done by myself jointly with others, I have made clear exactly what was done by others and what I have contributed myself.\\
\end{itemize}

Signed:\\
\rule[1em]{25em}{0.5pt} % This prints a line for the signature

Date:\\
\rule[1em]{25em}{0.5pt} % This prints a line to write the date
}

\clearpage % Start a new page

%----------------------------------------------------------------------------------------
%	QUOTATION PAGE
%----------------------------------------------------------------------------------------

\pagestyle{empty} % No headers or footers for the following pages

\null\vfill % Add some space to move the quote down the page a bit

\textit{``All truly great thoughts are conceived while walking."}
\begin{flushright}
Friedrich Nietzsche
\end{flushright}

\vfill\vfill\vfill\vfill\vfill\vfill\null % Add some space at the bottom to position the quote just right

\clearpage % Start a new page

%----------------------------------------------------------------------------------------
%	ABSTRACT PAGE
%----------------------------------------------------------------------------------------

\addtotoc{Abstract} % Add the "Abstract" page entry to the Contents

\abstract{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics

% The Thesis Abstract is written here (and usually kept to just this page). The page is kept centered vertically so can expand into the blank space above the title too\ldots

In this thesis we construct a method to create a temporal segmentation on time series data recorded during human activities by inertial sensors in smartphones.
Our assumption is that information about changes between activities can aid activity classification methods, since it adds information about sequential data points.
We explore current temporal segmentation methods and relate the problem to change detection.
A change in the performed activity results in a change in the data distribution of a time window.
Using a \acrlong{oc-svm} algorithm we model the data distribution in the shape of a high dimensional hypersphere.
A change in the distribution can then be discovered by analyzing the radius of the hypersphere, which will increase in such events.

We implement an algorithm to process real-world time series data and find changes in performed activities.
The data needs to be recorded in a continuous nature, with the transition periods between activities present.
Since common available data sets do not meet this requirement, we have recorded our own data set and made it publicly available.
A ground truth is provided both as manually annotated change points and video recordings of the activities.

The constructed method, \acrlong{ocs-hats}, is able to find changes between performed activities in an unknown environment.
}

\clearpage % Start a new page

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\setstretch{1.3} % Reset the line-spacing to 1.3 for body text (if it has changed)

\acknowledgements{\addtocontents{toc}{\vspace{0.8em}} % Add a gap in the Contents, for aesthetics

% The acknowledgements and the people to thank go here, don't forget to include your project advisor\ldots
Writing a thesis is hard.
It starts with the moment you are thinking about a subject and ends with the last part, where one should write the acknowledgments to any who deserves it and express thanks to those you owe it.
Besides all the authors who went before me in this subject and inspired me to continue in their line of research, there are a few people who I would like to say thanks to explicitly.

First, I am thankful for the attention from my supervisor at \textsc{Almende}, Anne van Rossum.
We had great discussions, in which I often needed to convince him of my ideas, which helped me with finding a domain and problem for my research.
I think it is great that \textsc{Almende} and \textsc{Sense} facilitate so much research for master students, in a very pleasant environment.
Everybody contributes to a atmosphere of research on new possibilities and applications, heavily relying on artificial intelligence.

Less frequent visited, but of no less importance, is my supervisor at the \textsc{Utrecht University}, Gerard Vreeswijk.
During our meetings he gave me great advise and feedback on my writing skills and techniques.
I am thankful for his feedback and guidance and allowing me to find a subject within my full interest.

Full immersion in a subject can narrow the field of vision.
Without reflecting on the subject from a small distance and a different point of view, the primary goals can easily become obfuscated.
The (for both sides) explanatory conversations with, amongst others, my partner Inge Dalm, my parents, my brother Floris, my sister Esmee, their partners, my housemate Yvo Zollner, those who are gathered under the heading of Lemnisco, those who call themselves `Vlassies', those who enjoy golden liquids, and those who are fond of lunacy, helped me understand my research problem at different levels.

Already mentioned, my parents (in the widest sense) have been of great support in any way.
It is their encouragement to explore the world and question everything that has led me to my passion for computer science, artificial intelligence,  and machine learning.

Finally, but certainly not least, I want to thank my partner Inge Dalm.
Without her and the time together I could not have completed this thesis.
I am very thankful for her and her vision on this matter.
She encouraged me during the process and bared with me during the harder moments.

With all the people mentioned above, and those who are not mentioned but know they should have, this research has become possible.
Without them, writing a thesis is even harder.
}
% \TODO{Write some nice stuff here}
\clearpage % Start a new page

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\pagestyle{fancy} % The page style headers have been "empty" all this time, now use the "fancy" headers as defined before to bring them back

\lhead{\emph{Contents}} % Set the left side page header to "Contents"
\tableofcontents % Write out the Table of Contents

\glsresetall

\lhead{\emph{List of Figures}} % Set the left side page header to "List of Figures"
\listoffigures % Write out the List of Figures

\glsresetall

\lhead{\emph{List of Tables}} % Set the left side page header to "List of Tables"
\listoftables % Write out the List of Tables

\glsresetall

% %----------------------------------------------------------------------------------------
% %	ABBREVIATIONS
% %----------------------------------------------------------------------------------------

% \clearpage % Start a new page

% \setstretch{1.5} % Set the line spacing to 1.5, this makes the following tables easier to read

% \lhead{\emph{Abbreviations}} % Set the left side page header to "Abbreviations"
% \listofsymbols{ll} % Include a list of Abbreviations (a table of two columns)
% {
% \textbf{LAH} & \textbf{L}ist \textbf{A}bbreviations \textbf{H}ere \\
% %\textbf{Acronym} & \textbf{W}hat (it) \textbf{S}tands \textbf{F}or \\
% \textbf{PCA} & \textbf{P}rinciple \textbf{C}omponent \textbf{A}nalysis
% }

%----------------------------------------------------------------------------------------
% GLOSSARIES
%----------------------------------------------------------------------------------------

\clearpage

\lhead{\emph{Acronyms}}

\printglossaries

% %----------------------------------------------------------------------------------------
% %	PHYSICAL CONSTANTS/OTHER DEFINITIONS
% %----------------------------------------------------------------------------------------

% \clearpage % Start a new page

% \lhead{\emph{Physical Constants}} % Set the left side page header to "Physical Constants"

% \listofconstants{lrcl} % Include a list of Physical Constants (a four column table)
% {
% Speed of Light & $c$ & $=$ & $2.997\ 924\ 58\times10^{8}\ \mbox{ms}^{-\mbox{s}}$ (exact)\\
% % Constant Name & Symbol & = & Constant Value (with units) \\
% }

%----------------------------------------------------------------------------------------
%	SYMBOLS
%----------------------------------------------------------------------------------------

% \clearpage % Start a new page

% \lhead{\emph{Symbols}} % Set the left side page header to "Symbols"

% \listofnomenclature{lll} % Include a list of Symbols (a three column table)
% {
% $a$ & distance & m \\
% $P$ & power & W (Js$^{-1}$) \\
% % Symbol & Name & Unit \\

% & & \\ % Gap to separate the Roman symbols from the Greek

% $\omega$ & angular frequency & rads$^{-1}$ \\
% % Symbol & Name & Unit \\
% }

%----------------------------------------------------------------------------------------
%	DEDICATION
%----------------------------------------------------------------------------------------

% \clearpage
% \setstretch{1.3} % Return the line spacing back to 1.3

% \pagestyle{empty} % Page style needs to be empty for this page

% \dedicatory{For/Dedicated to/To my\ldots} % Dedication text

% \addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{fancy} % Return the page headers back to the "fancy" style

% Include the chapters of the thesis as separate files from the Chapters folder
% Uncomment the lines as you write the chapters

\input{./Chapters/Chapter1}
\input{./Chapters/Chapter2}
\input{./Chapters/Chapter3}
\input{./Chapters/Chapter4}
\input{./Chapters/Chapter5}
\input{./Chapters/Chapter6}
\input{./Chapters/Chapter7}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

\appendix % Cue to tell LaTeX that the following 'chapters' are Appendices

% Include the appendices of the thesis as separate files from the Appendices folder
% Uncomment the lines as you write the Appendices

\input{./Appendices/AppendixA-Results}
% \input{./Appendices/AppendixA-Summaries}
% \input{./Appendices/AppendixB-Camci}
% \input{./Appendices/AppendixC}
% \input{./Appendices/AppendixD-Planning}
% \input{./Appendices/AppendixE-Data-structure}
% \input{./Appendices/AppendixF-final}

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

\backmatter

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\label{Bibliography}

\lhead{\emph{Bibliography}} % Change the page header to say "Bibliography"

% \bibliographystyle{unsrtnat} % Use the "unsrtnat" BibTeX style for formatting the Bibliography
\bibliographystyle{plainnat} % Use the "unsrtnat" BibTeX style for formatting the Bibliography

\bibliography{Bibliography} % The references (bibliography) information are stored in the file named "Bibliography.bib"

\end{document}