### 2014-01-20 Chapters/Chapter6.tex

% !TEX root = ../main.tex
% Chapter 6

\chapter{Real-world results}

\label{Chapter6} % For referencing the chapter elsewhere, use~\ref{Chapter6}

\lhead{Chapter 6. \emph{Real-world results}} % This is for the header on each page - perhaps a shortened title

In this chapter we discuss the real-world data sets used for this research and the performance of our proposed algorithm on these data sets.
In \Cref{sec:data_sets} we discuss available data sets and argue that these are inappropriate for our research.
It is followed by an overview of the setup we have created to record our own continuous data sets in an unknown environment, both for in- and outdoor activities.
The series of activities performed are discussed in \Cref{sec:data_gathering}.
In that section we show typical plots of the inertial sensor data and stills from the video recordings.
That will illustrate the setting in which the activities are performed and the characteristics of continuous data.
The results of the algorithm are discussed in \Cref{sec:real_world_results}.
Finally, we provide a list of possible improvements in \Cref{sec:possible_improvements}.

The results are measured with both quantitative measures, as introduced in \Cref{sec:artificial_data_quality_metrics}, and qualitative claims on the performance of the algorithm applied to the real-world data.
Since the data sets used are real-world recordings, we will put emphasis on the latter type of quality measures \ie the qualitative claims.

\input{./Chapters/Chapter6/data_sets}
\input{./Chapters/Chapter6/data_gathering}
\input{./Chapters/Chapter6/results}